A user-extensible and adaptable parser architecture

### 2008-07-30 doc/logparser-results.tex
Further corrections from Carl. Added paragraph about pathological rules. Some cleanups and XXX items dealt with.

CPU         & One dual core 2.40GHz Intel(R) Core(TM)2 CPU, \\
            & with 32KB L1 cache and 4MB L2 cache. \\
RAM         & 2GB 667 MHz DDR RAM. \\
Hard disk   & One XXX 232GB SATA disk. \\

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of lines graph} shows the
parsing time in seconds, file size in MB, and tens of thousands of log
lines for each of the \numberOFlogFILES{} log files.  All three lines run
roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor} which plots
both the ratio of file size vs parsing time, and the ratio of number of log
lines vs parsing time (higher is better in both cases); \tableref{parsing
time vs file size vs number lines factor table} shows the same ratios for
different groups of log files.  As the reader can see the ratios are quite
tightly banded, providing empirical evidence that the algorithm scales
linearly.  The ratio increases (i.e.\ improves) for log files 22 and 62--68
despite their larger than average size (shown in \graphref{parsing time vs
file size vs number of lines graph}); this strange behaviour where larger
log files are parsed more efficiently than smaller is explained fully in
\sectionref{Explaining the peaks in log file size}.

>>>>

CPU         & One dual core 2.40GHz Intel(R) Core(TM)2 CPU, \\
            & with 32KB L1 cache and 4MB L2 cache. \\
RAM         & 2GB 667 MHz DDR RAM\@. \\
Hard disk   & One XXX 232GB SATA disk. \\

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of lines graph} shows the
parsing time in seconds, file size in MB, and tens of thousands of log
lines for each of the \numberOFlogFILES{} log files.  All three lines run
roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor} which plots
both the ratio of file size vs parsing time, and the ratio of number of log
lines vs parsing time (higher is better in both cases); \tableref{parsing
time vs file size vs number lines factor table} shows the same ratios for
different groups of log files.  The ratios are quite tightly banded,
providing empirical evidence that the algorithm scales linearly.  The ratio
increases (i.e.\ improves) for log files 22 and 62--68 despite their larger
than average size (shown in \graphref{parsing time vs file size vs number
of lines graph}); this is explained fully in \sectionref{Explaining the
peaks in log file size}.

### 2008-07-30 papers/sgai-2008/paper.tex
Further corrections from Carl. Added paragraph about pathological rules. Some cleanups and XXX items dealt with.

% XXX FIGURE OUT HOW TO DO THIS AUTOMATICALLY
% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    68.04\%%
}

@@

Rules categorise inputs, specifying both the data to be extracted from each
log line and the action to invoke.  The Postfix log parser stores the rules
in the same SQL database that the results are stored in, removing any doubt
as to which set of rules was used to produce a set of results; other
implementations are free to store the rules in whatever fashion suits their
needs.

XXX ADD SOMETHING ABOUT IMPLEMENTATIONS BEING FREE TO ADD OR REMOVE
ATTRIBUTES\@.

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
SLCT~\cite{slct-paper} to aid in producing regexes to parse previously
unparsed inputs.  Storing the rules separately allows other approaches to
rule management to be used, e.g.\ instead of manually adding new rules when
unparsed log lines are seen, machine learning techniques could be used to
automatically generate new rules, based on the existing rules and input
lines.  If this approach was taken the choice of machine learning technique
would be constrained by the size of typical data sets: the
\numberOFlogFILES{} log files used to generate statistics contain
\numberOFlogLINEShuman{} log lines.  Techniques which require the full data
set when training would be impractical; Instance Based Learning
(IBL)~\cite{instance-based-learning} techniques which automatically
determine which inputs from the training set are valuable and which can be
discarded would probably reduce the data required to a manageable size.
This avenue of research and development has not been pursued by the author,
but could easily be independently undertaken.

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins, and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  The algorithm provides
a facility for specifying the order of overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches they had parsing the previous log file.  Negative
priorities may be useful for catchall rules.

@@

XXX TIMING INFORMATION\@.
\newpage

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and tens of thousands of log lines for
each of the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor} which plots both the ratio of file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is better in both cases).  As the reader can see the ratios are quite
tightly banded across the graph, showing that the algorithm scales
linearly.  The ratio increases (i.e.\ improves) for log files 22 and 62--68
despite their larger than average size (shown in \graphref{Parsing time,
file size, and number of lines}); this strange behaviour where larger log
files are parsed more efficiently is explained in \sectionref{Why are there
dips in the graphs?}.
\showgraph{build/plot-normal-filesize-numlines}{Parsing time, file size,
and number of lines}{Parsing time, file size, and number of lines}
\showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file size
and number of lines to parsing time}{parsing time vs file size vs number
lines factor}

>>>>

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    68.04\%%
}

@@

Rules categorise inputs, specifying both the data to be extracted from each
log line and the action to invoke.  The Postfix parser stores the rules in
the same SQL database that the results are stored in, removing any doubt as
to which set of rules was used to produce a set of results; other
implementations are free to store the rules in whatever fashion suits their
needs.  Similarly implementations are free to require additional attributes
in their rules; the framework requires \texttt{regex} and \texttt{action}
attributes, to determine whether a rule matches the current input line, and
which action to invoke if it does.  Rules used by the Postfix parser
define the following attributes: XXX DO I NEED TO LIST ALL THE
ATTRIBUTES\@?

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
SLCT~\cite{slct-paper} to aid in producing regexes to parse previously
unparsed inputs.  Storing the rules separately allows other approaches to
rule management to be used, e.g.\ instead of manually adding new rules when
unparsed log lines are seen, machine learning techniques could be used to
automatically generate new rules, based on the existing rules and input
lines.  If this approach was taken the choice of machine learning technique
would be constrained by the size of typical data sets: the
\numberOFlogFILES{} log files used to generate statistics contain
\numberOFlogLINEShuman{} log lines.  Techniques which require the full data
set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques which automatically
determine which inputs from the training set are valuable and which inputs
can be discarded would probably reduce the data required to a manageable
size.  This avenue of research and development has not been pursued by the
author, but could easily be independently undertaken.

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins, and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  The algorithm provides
a facility for specifying the order of overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches they had parsing the previous log file.
Negative priorities may be useful for catchall rules.  XXX DETECTING
OVERLAP\@.

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
large amounts of CPU time during matching.  Defining a regex which matches
every input line is trivial: \verb!/^/! will match the start of the input
line; every input line is guaranteed to start, so this regex would match
every input line.  In general a regex will not consume a large amount of
CPU time when the match is successful; usually CPU time is consumed by a
regex with a lot of alteration and variable quantifiers failing to match.
This topic is beyond the scope of this paper, see
\cite{mastering-regular-expressions} for in-depth discussion.

@@

XXX TIMING INFORMATION\@.

XXX TIMING INFORMATION\@.

XXX TIMING INFORMATION\@.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and tens of thousands of log lines for
each of the \numberOFlogFILES{} log files.  All three lines run roughly in
parallel, giving the impression that the algorithm scales linearly with
input size.  This impression is borne out by \graphref{parsing time vs file
size vs number lines factor} which plots both the ratio of file size vs
parsing time, and the ratio of number of log lines vs parsing time (higher
is better in both cases).  The ratios are quite tightly banded across the
graph, showing that the algorithm scales linearly.  The ratio increases
(i.e.\ improves) for log files 22 and 62--68 despite their larger than
average size (shown in \graphref{Parsing time, file size, and number of
lines}); this is explained in \sectionref{Why are there dips in the
graphs?}.  \showgraph{build/plot-normal-filesize-numlines}{Parsing time,
file size, and number of lines}{Parsing time, file size, and number of
lines} \showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file
size and number of lines to parsing time}{parsing time vs file size vs
number lines factor}
